Cross-disciplinary assessment of statistical guidance provided to authors by influential academic journals

### 2021-09-02 analysis/preliminaryAnalysis.Rmd
preliminary analysis 1st full draft

---
title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
author: "Analysis performed by: Maia Salholz-Hillel & Tom Hardwicke"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    number_sections: true
    global_numbering: true
editor_options: 
  chunk_output_type: console
---

@@

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(lemon) # to have axes across all graph facets
library(ggthemr) # for ggplot theme
library(ggrepel) # for ggplot text labels
library(scales) # to wrap text on axis labels
```

```{r load_functions}
# load custom functions
source(here('analysis','functions.R'))
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves preprocessed files
source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files output by the preprocessing performed above
load(here('data','processed','d_all.rds'))
load(here('data','processed','lookup_external_guidance.rds'))
```

@@

# Introduction
<br>
We have now completed data collection for the first phase of this project. We extracted author-facing statistical guidance provided by the top-15 journals (by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals). The preregistered study protocol is available [here](https://osf.io/cz9g3/). Below we present a preliminary analysis of the phase one data and highlight some issues to discuss as a team.

@@

> **Operational definition of statistical guidance**: any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.

We recorded whether each journal:

@@

* We coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.

**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not extracted guidance from those sources*. We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.

@@

@@

**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals). Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to. For example, the Cell journals' guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset. Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.

@@

@@

  # Recode NAs to FALSE
  mutate(across(
    everything()
    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
    , ~replace_na(., FALSE) )) %>%

  select(esi_field, starts_with("has_")) %>% 
  group_by(esi_field) %>%
  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
  adorn_totals(name = 'OVERALL') %>%
  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
  mutate(prop = n/N, percent = round(prop*100, 0)) %>%
  mutate(esi_abbr = case_when( # create abbreviated field labels
    esi_field == "OVERALL" ~ "OVERALL",
    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
    esi_field == "CLINICAL MEDICINE" ~ "MED",           
    esi_field == "COMPUTER SCIENCE" ~ "COMSCI",             
    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
    esi_field == "ENGINEERING" ~ "ENGIN",                  
    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
    esi_field == "GEOSCIENCES" ~ "GEO",                  
    esi_field == "IMMUNOLOGY" ~ "IMMUN",                   
    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
    esi_field == "MATHEMATICS" ~ "MATH",                
    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
    esi_field == "Multidisciplinary" ~ "MULTI",            
    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
    esi_field == "PHYSICS" ~ "PHYS",                     
    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCSCI",     
    esi_field == "SPACE SCIENCE" ~ "SPACE"          
  )) %>%
  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')')) %>%
  mutate(var_display = case_when( # create display versions of topic names
    variable == "guidance" ~ "Provides any statistical guidance",
    variable == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    variable == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    variable == "external_guidance" ~ "Refers to external statistical guidance",
    variable == "p_value" ~ "p values",
    variable == "significance" ~ "statistical significance",
    variable == "null_hypo" ~ "null hypotheses",
    variable == "sample_size" ~ "sample size justification",
    variable == "conf_int" ~ "confidence intervals",
    variable == "effect_size" ~ "effect sizes",
    variable == "multi_compare" ~ "multiple comparisons",
    variable == "subgroup" ~ "subgroup analyses",
    variable == "baseline_covar" ~ "baseline covariates",
    variable == "non_param" ~ "non-parametric tests",
    variable == "sensitivity" ~ "sensitivity analyses",
    variable == "model_assume" ~ "checking model assumptions",
    variable == "exclusion" ~ "data exclusions",
    variable == "outliers" ~ "handling outliers",
    variable == "missing" ~ "handling missing data",
    variable == "one_sided" ~ "one-sided tests",
    variable == "bayes" ~ "Bayesian statistics",
    variable == "secondary" ~ "secondary outcomes",
    variable == "prespecify" ~ "prespecification of analyses",
    variable == "cat_continuous" ~ "categorisation of continuous data",
    variable == "publisher_guidance" ~ "Shares publisher guidance"
  ))

@@

# select only data on general any/internal/external statistical guidance classifications 
sum_tab_general <- sum_tab %>%
  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')),
         var_display = factor(var_display, levels = c("Provides any statistical guidance", "Provides journal-specific statistical guidance", "Has a dedicated statistical guidance section", "Refers to external statistical guidance"))) # set variable order

# select only data on topics
sum_tab_topics <- sum_tab %>%
  filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance', 'publisher_guidance')) %>% # select only data on topics
  rename(topic = variable, topic_display = var_display)

@@

Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that (a) offered any statistical guidance; (b) had a dedicated statistical guidance section. For tabular data, see Appendix \@ref(app).

@@

You can see that around half (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) of journals offered any statistical guidance. Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers). Just over quarter of journals had a dedicated statistical guidance section of their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%). In two fields (Computer Science and Maths), no journals offered any statistical guidance. Journals in health-related fields were more likely to offer statistical guidance and have dedicated statistical guidance sections. Notably, 100% of the journals in clinical medicine offered some statistical guidance.

<br>

```{r statguidefig, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond). Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 6}
# extract overall data
sum_tab_general_overall <- sum_tab_general %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  filter(esi_field == "OVERALL")

sum_tab_general %>%
  filter(esi_field != "OVERALL") %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 15,
    #nudge_x = 5,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```

<br><br>

Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty pre-specified statistical topics. For tabular data, see the Appendix (Section \@ref(app)). Please see Section \@ref(issues) (issues) at the end of this document for an important note regarding the topic "prespecification of analyses".

<br>

```{r stattopicsfig, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond). For presentational purposes, disciplines offering no guidance on individual topics are shown, but not labelled. Two scientific disciplines in which no journals offered any statistical guidance at all are not shown. Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 28}
# extract overall data
sum_tab_topics_overall <- sum_tab_topics %>%
  filter(esi_field == "OVERALL")

@@

<br>

You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuus data). Journals in health-related fields, particularly clinical medicine, are often more likely to mention these topics than journals in other fields.

@@

<br>

@@

<br>

```{r topicshist, fig.cap = 'Histogram showing how many of the twenty statistical topics were mentioned by each journal'}
n_topics %>%
  ggplot(aes(x=topics_mentioned)) +
  geom_histogram()
```

<br>

In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty topics were mentioned by each journal. Journals that shared publisher guidance are represented by one row (per publisher). Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown.

<br>

```{r}
my_palette <- c("#56B4E9", "#E69F00")

@@

plt <- d_internal_guidance %>%
  mutate(journal = factor(journal, levels = journals_topics_mentioned_ordered)) %>% # order journals by n topics mentioned
  select(journal, starts_with("has_") & (!contains("guidance"))) %>%
  mutate(journal = fct_recode(journal,"PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA")) %>% # shorten a particularly long journal name
  pivot_longer(
    -journal,
    names_to = "topic",
    names_prefix = 'has_',
    values_to = "outcome",
    values_drop_na = FALSE
  ) %>%
  mutate(outcome = case_when(
    outcome ~ "Mentioned",
    !outcome ~ "Not mentioned",
    TRUE ~ "Not applicable (No internal guidance)"
  )) %>%
  mutate(topic_display = case_when( # create display versions of topic names
    topic == "guidance" ~ "Provides any statistical guidance",
    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    topic == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    topic == "external_guidance" ~ "Refers to external statistical guidance",
    topic == "p_value" ~ "p values",
    topic == "significance" ~ "statistical significance",
    topic == "null_hypo" ~ "null hypotheses",
    topic == "sample_size" ~ "sample size justification",
    topic == "conf_int" ~ "confidence intervals",
    topic == "effect_size" ~ "effect sizes",
    topic == "multi_compare" ~ "multiple comparisons",
    topic == "subgroup" ~ "subgroup analyses",
    topic == "baseline_covar" ~ "baseline covariates",
    topic == "non_param" ~ "non-parametric tests",
    topic == "sensitivity" ~ "sensitivity analyses",
    topic == "model_assume" ~ "checking model assumptions",
    topic == "exclusion" ~ "data exclusions",
    topic == "outliers" ~ "handling outliers",
    topic == "missing" ~ "handling missing data",
    topic == "one_sided" ~ "one-sided tests",
    topic == "bayes" ~ "Bayesian statistics",
    topic == "secondary" ~ "secondary outcomes",
    topic == "prespecify" ~ "prespecification of analyses",
    topic == "cat_continuous" ~ "categorisation of continuous data",
    topic == "publisher_guidance" ~ "Shares publisher guidance"
  )) %>%
  ggplot(aes(x = topic_display, y = journal)) +
    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
    scale_fill_manual(
      values = my_palette,
      # labels = c('Not mentioned', 'Unclear', 'Mentioned')
    ) +
    scale_x_discrete(position = 'top') +
    scale_y_discrete(labels = wrap_format(25)) + # wrap text on y axis
    xlab("Statistical topics") +
    theme(plot.margin = margin(0, 50, 0, 0),
          axis.ticks = element_blank(),
          axis.title.y =element_blank(),
          axis.line = element_blank(),
          panel.grid.major=element_line(colour=NA),
          panel.grid.minor=element_line(colour=NA),
          legend.position = 'top',
          axis.text.x  = element_text(angle = 45, hjust = 0),
          strip.background = element_rect(
            color="black", fill="white", size=0, linetype="solid"
          ),
          strip.text.y = element_text(
            size = 8, color = "black", face = 'bold'
          ))

@@

```{r guidancemosaic, fig.cap = "Grid diagram showing whether each journal provdied guidance on each of twenty prespecified statistical topics", fig.height = 35, fig.width = 10}
plt
```

@@

To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(fig:guidanceexamples) below.

```{r guidanceexamples}
guidance_examples_tbl %>% kable()
```

Look at external guidance

```{r tbl_external_guidance}
lookup_external_guidance %>% 
  knitr::kable()
```

External guidance types

```{r tbl_external_guidance_types}
lookup_external_guidance %>% 
  count(guidance_type) %>% 
  kable()
```

>>>>

---
title: "Cross-disciplinary assessment of statistical guidance provided to authors by influential academic journals"
description: |
  Preliminary analysis for study one and study two.
author:
  - name: Maia Salholz-Hillel & Tom Hardwicke
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
---

@@

```{r load_packages}
library(knitr) # for literate programming
library(tidyverse) # for data munging
library(janitor) # for data munging
library(tidylog) # inline code feedback
library(here) # for finding files
library(assertr) # for testing
library(lemon) # to have axes across all graph facets
library(ggthemr) # for ggplot theme
library(ggrepel) # for ggplot text labels
library(scales) # to wrap text on axis labels
library(rmarkdown) # for paged table
```

```{r load_functions}
source(here('analysis','functions.R')) # load custom functions
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves processed data files
#source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files
load(here('data','processed','d_all.rds'))
load(here('data','processed','lookup_external_guidance.rds'))
```

@@

# Introduction

We have now completed dual-coded data collection for the first study of this project and single-coded data collection for the second study of this project. The preregistered study protocol is available [here](https://osf.io/cz9g3/). Below we present a preliminary analysis of the data and highlight some issues to discuss as a team.

@@

## Study one

In November, 2019, we made copies of any documentation (e.g., instructions to authors) that provided author-facing statistical guidance provided by the top-15 journals (ranked by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).

**Operational definition of statistical guidance**

> any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.

We then examined this documentation and recorded whether each journal:

@@

* We also recorded whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.

@@

**Note:** When journals referred to guidance in external sources (typically reporting guidelines), we have recorded the names of the sources, but *we have not extracted guidance from those sources*.

@@

**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals). An additional journal - Scientific Data - shares guidance with other Nature journals, but also has its own specific guidance. Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to. For example, the `r publisher_cell_n` journals published by Cell inherit Cell's "STAR Methods" guidance, so this guidance appears `r publisher_cell_n` times in our dataset.

## Study two

For study two, we planned to more closely examine the guidance provided on five specific topics:

- p-values
- statistical significance
- confidence intervals
- effect sizes
- sample size justification

We prespecified some information we wanted to extract about these topics; however, when I (Tom) examined the verbatim guidance, I found our prespecified questions and response options to be unsuitable (ambiguous, badly worded, etc). So in an entirely data-dependent and exploratory manner, I instead coded the level of endorsement for each topic/method according to four categories:

1. Explicit endorsement - the journal advises or instructs authors to use this method whenever possible/appropriate.
2. Implicit endorsement - the journal provides advice on the method, implying endorsement, but does not explictly advise that the method should be used.
3. Implicit opposition - the journal advises that they would prefer the method is not used, but does not explicitly rule it out.
4. Explicit opposition - the journal advises that the method should not be used.

This was admittedly quite subjective and I had to 'read between the lines' a bit in order to assign guidance to these categories.

I also categorised reporting advice for three topics as below:

p-values:

1. Report exact p-values
2. Report exact p-values unless very small
3. Report p-values relative to thresholds (e.g., p < .05)
4. No guidance

Statistical significance:

1. Report alpha
2. Use specific alpha (e.g., a = .05)
3. No guidance

Confidence intervals:

1. Report confidence level
2. Use specific confidence level (e.g., 95%)
3. No guidance

So far, the data coding for study two has only been performed by me (Tom), though we originally planned for dual-coding (see issues section at the end of this document).

@@

## Study one

@@

  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  select(esi_field, starts_with("has_")) %>% 
  group_by(esi_field) %>%
  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
  adorn_totals(name = 'OVERALL') %>%
  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
  mutate(prop = n/N, percent = round(prop*100, 0)) %>%
  mutate(esi_abbr = case_when( # create abbreviated field labels to use in figures
    esi_field == "OVERALL" ~ "OVERALL",
    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
    esi_field == "CLINICAL MEDICINE" ~ "MED",           
    esi_field == "COMPUTER SCIENCE" ~ "COMSCI",             
    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
    esi_field == "ENGINEERING" ~ "ENGIN",                  
    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
    esi_field == "GEOSCIENCES" ~ "GEO",                  
    esi_field == "IMMUNOLOGY" ~ "IMMUN",                   
    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
    esi_field == "MATHEMATICS" ~ "MATH",                
    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
    esi_field == "Multidisciplinary" ~ "MULTI",            
    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
    esi_field == "PHYSICS" ~ "PHYS",                     
    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCSCI",     
    esi_field == "SPACE SCIENCE" ~ "SPACE"          
  )) %>%
  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')')) %>%
  mutate(var_display = case_when( # create display versions of topic names
    variable == "guidance" ~ "Provides any statistical guidance",
    variable == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    variable == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    variable == "external_guidance" ~ "Refers to external statistical guidance",
    variable == "p_value" ~ "p values",
    variable == "significance" ~ "statistical significance",
    variable == "null_hypo" ~ "null hypotheses",
    variable == "sample_size" ~ "sample size justification",
    variable == "conf_int" ~ "confidence intervals",
    variable == "effect_size" ~ "effect sizes",
    variable == "multi_compare" ~ "multiple comparisons",
    variable == "subgroup" ~ "subgroup analyses",
    variable == "baseline_covar" ~ "baseline covariates",
    variable == "non_param" ~ "non-parametric tests",
    variable == "sensitivity" ~ "sensitivity analyses",
    variable == "model_assume" ~ "checking model assumptions",
    variable == "exclusion" ~ "data exclusions",
    variable == "outliers" ~ "handling outliers",
    variable == "missing" ~ "handling missing data",
    variable == "one_sided" ~ "one-sided tests",
    variable == "bayes" ~ "Bayesian statistics",
    variable == "secondary" ~ "secondary outcomes",
    variable == "prespecify" ~ "prespecification of analyses",
    variable == "cat_continuous" ~ "categorisation of continuous data",
    variable == "publisher_guidance" ~ "Shares publisher guidance"
  ))

@@

# select only variables that address the general any/internal/external statistical guidance classifications 
sum_tab_general <- sum_tab %>%
  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')),
         var_display = factor(var_display, levels = c("Provides any statistical guidance", "Provides journal-specific statistical guidance", "Has a dedicated statistical guidance section", "Refers to external statistical guidance"))) # set variable order

# select only variables on specific topics
sum_tab_topics <- sum_tab %>%
  filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance', 'publisher_guidance')) %>% # select only data on topics
  rename(topic = variable, topic_display = var_display)

@@

Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that:

(a) offered any statistical guidance; 
(b) had a dedicated statistical guidance section.

For tabular data, see the table in the Appendix.

@@

You can see that around half (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) of journals offered any statistical guidance. Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers). Just over quarter of journals had a dedicated statistical guidance section in their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%). In two fields (Computer Science and Maths), no journals offered any statistical guidance. Journals in health-related fields were more likely to offer statistical guidance and have dedicated statistical guidance sections. Notably, 100% of the journals in clinical medicine offered some statistical guidance.

```{r}
# extract overall data
sum_tab_general_overall <- sum_tab_general %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  filter(esi_field == "OVERALL")
```

```{r statguidefig, layout="l-page", fig.width = 10, fig.height = 6, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond)."}
sum_tab_general %>%
  filter(esi_field != "OVERALL") %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 15,
    #nudge_x = 5,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```
Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty statistical topics we pre-selected. For tabular data, see the table in the Appendix. Please see the issues section at the end of this document for an important note regarding the topic "prespecification of analyses".

You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuous data). Journals in health-related fields, particularly clinical medicine, are often more likely to mention this collection of topics than journals in other fields.

```{r stattopicsfig, layout="l-page", fig.width = 10, fig.height = 28, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond). For presentational purposes, two scientific disciplines in which no journals offered any statistical guidance at all are not shown. Additionally, disciplines offering no guidance on individual topics are shown, but not labelled. Graphs are ordered from left to right and top to bottom in order of highest proportion offering topic guidance overall across scientific disciplines."}
# extract overall data
sum_tab_topics_overall <- sum_tab_topics %>%
  filter(esi_field == "OVERALL")

@@

@@

@@

```{r topicshist, fig.cap = 'Histogram showing how many of the twenty statistical topics were mentioned by each journal'}
n_topics %>%
  ggplot(aes(x=topics_mentioned)) +
  geom_histogram()
```
In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty preselected topics were mentioned by each journal. Journals that shared publisher guidance are represented by one row (per publisher). Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown.

```{r}
my_palette <- c("#95D840FF", "#404788FF")

@@

plt <- d_internal_guidance %>%
  mutate(journal = factor(journal, levels = journals_topics_mentioned_ordered)) %>% # order journals by n topics mentioned
  select(journal, starts_with("has_") & (!contains("guidance"))) %>%
  mutate(journal = fct_recode(journal,"PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA")) %>% # shorten a particularly long journal name
  pivot_longer(
    -journal,
    names_to = "topic",
    names_prefix = 'has_',
    values_to = "outcome",
    values_drop_na = FALSE
  ) %>%
  mutate(outcome = case_when(
    outcome ~ "Mentioned",
    !outcome ~ "Not mentioned",
    TRUE ~ "Not applicable (No internal guidance)"
  )) %>%
  mutate(topic_display = case_when( # create display versions of topic names
    topic == "guidance" ~ "Provides any statistical guidance",
    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    topic == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    topic == "external_guidance" ~ "Refers to external statistical guidance",
    topic == "p_value" ~ "p values",
    topic == "significance" ~ "statistical significance",
    topic == "null_hypo" ~ "null hypotheses",
    topic == "sample_size" ~ "sample size justification",
    topic == "conf_int" ~ "confidence intervals",
    topic == "effect_size" ~ "effect sizes",
    topic == "multi_compare" ~ "multiple comparisons",
    topic == "subgroup" ~ "subgroup analyses",
    topic == "baseline_covar" ~ "baseline covariates",
    topic == "non_param" ~ "non-parametric tests",
    topic == "sensitivity" ~ "sensitivity analyses",
    topic == "model_assume" ~ "checking model assumptions",
    topic == "exclusion" ~ "data exclusions",
    topic == "outliers" ~ "handling outliers",
    topic == "missing" ~ "handling missing data",
    topic == "one_sided" ~ "one-sided tests",
    topic == "bayes" ~ "Bayesian statistics",
    topic == "secondary" ~ "secondary outcomes",
    topic == "prespecify" ~ "prespecification of analyses",
    topic == "cat_continuous" ~ "categorisation of continuous data",
    topic == "publisher_guidance" ~ "Shares publisher guidance"
  )) %>%
  mutate(topic_display = factor(topic_display, levels = ordered_topics)) %>% # order topics by n topics mentioned
  ggplot(aes(x = topic_display, y = journal)) +
    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
    scale_fill_manual(
      values = my_palette,
      # labels = c('Not mentioned', 'Unclear', 'Mentioned')
    ) +
    scale_x_discrete(position = 'top') +
    scale_y_discrete(labels = wrap_format(25)) + # wrap text on y axis
    xlab("Statistical topics") +
    theme(plot.margin = margin(0, 50, 0, 0),
          axis.ticks = element_blank(),
          axis.title.y =element_blank(),
          axis.line = element_blank(),
          panel.grid.major=element_line(colour=NA),
          panel.grid.minor=element_line(colour=NA),
          legend.position = 'top',
          axis.text.x  = element_text(angle = 45, hjust = 0),
          strip.background = element_rect(
            color="black", fill="white", size=0, linetype="solid"
          ),
          strip.text.y = element_text(
            size = 8, color = "black", face = 'bold'
          ))

@@

```{r guidancemosaic, layout="l-page", fig.cap = "Grid diagram showing whether each journal provided guidance on each of twenty preselected statistical topics. Topics are ordered from most (left) to least (right) mentioned. Journals are ordered from most (top) to least (bottom) preselected topics mentioned.", fig.height = 35, fig.width = 10}
plt
```

@@

To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(tab:guidanceexamples) below.

```{r guidanceexamples}
guidance_examples_tbl %>% kable(caption = "Randomly selected verbatim statistical guidance for each of 20 preselected statistical topics.")
```

`r d_all %>% filter(has_external_guidance == T) %>% nrow()` journals referred authors to statistical guidance in an external source (reporting guidelines or other sources like academic papers or websites). Of these, `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == F) %>% pull(n)` only referred to statistical guidance in external sources, whereas the other `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == T) %>% pull(n)` also provided their own guidance in additional to referring to external sources.

In the histogram (Figure \@ref(fig:repguidehist)) below, you can see how frequently journals referred to different reporting guidelines. Then in Table \@ref(tab:tbl-external-other-guidance) you can see all of the other external guidelines (academic papers and websites) that journals referred to. None of these external sources that have been examined by us in detail.
